Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-07-30 manuscript/revision2/revision2.tex
revision2: Coauthor line edits from SS+FM+RM+DL+AH+FF

\authorcontributions{T.E.J. designed research. D.B.L. contributed plant
material and resources. J.B., D.B.L., and T.E.J. designed and executed
field experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W.,
and T.E.J. hosted field experiments. A.H.M., L.Z., and S.A.S. conducted
statistical and computational analyses. The manuscript was written by
A.H.M. with contributions from all authors.}

@@

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depends on genetic effects, environmental
(e.g., weather) cues, and possibly their interaction. Quantifying these
GxWeather interactions can aid prediction and manipulation of
phenological timings. Here, we map GxWeather effects on phenological
timings in two highly divergent switchgrass (\emph{Panicum virgatum})
populations using repeated plantings of cloned individuals from these
populations at eight sites across the central United States. Some
genetic effects covary with interpretable weather-based cues. For
example, in the Gulf population, 65\% of genetic effects on the timing
of vegetative growth covary with daylength 14 days prior to green-up
date, and 33\% of genetic effects on the timing of flowering covary with
cumulative rainfall in the seven days prior to flowering. However, most
variation in genetic effects is site-specific and can not be attributed
to variation in the studied weather variables. We demonstrate that we
can identify genetic variation with GxWeather and assign these loci to
specific weather-based cues or other patterns. Breeding for particular
alleles at the GxWeather loci we identified could change flowering
responsiveness in a photoperiod or rainfall-specific way. More broadly,
our approach proposes a refined, powerful characterization of
genotype-by-environment interactions in any species phenotyped in
multiple environments.
\end{abstract}

@@

Plant phenological timings are major components of plant fitness
affected by multiple external environmental cues (e.g.~degree of winter
chilling, day length, temperature, and water availability) that signal
existing or upcoming growing conditions (1--4). Genetic responses to
environmental cues determine the speed, timing, and energy apportioned
to vegetative and reproductive growth and shape the individual's
physiological responses, lifespan, and lifetime production of viable
seed. Day length (or photoperiod) is one of the most predictable
environmental cues, and genetic sensitivity to photoperiod protects
plants from potentially fatal consequences of phenological responses to
temperature cues at the ``wrong'' time of year. However, the usefulness
of specific environmental cues depends on both features of the
environment, such as cue predictability and relevance, and the species'
adaptive strategies (5). Species with wide natural distributions can
have multiple distinct environmentally cued phenological responses: for
example, populations of sunflower (\emph{Helianthus annuus}) exhibit
day-neutral, facultative short day, and facultative long-day flowering
responses, which vary with their environments (6, 7). Distinct genetic
responses in different environments are known as genotype-by-environment
interactions, or GxE.

@@

Previous research suggests that switchgrass phenological timings should
have GxWeather and that these timings could differ by genetic
subpopulation. Switchgrass is considered a short-day plant with
reproductive development strongly linked to day of the year (35).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (36, 37). We previously found divergent
Midwest and Gulf genetic subpopulations of switchgrass with distinct
sets of environmental adaptations, in that both populations had distinct
genetic variation associated with each of two fitness proxies, biomass
and overwinter survival (38). The Midwest genetic subpopulation is
primarily composed of individuals from the well-studied upland
switchgrass ecotype (39, 40), while the Gulf subpopulation has
individuals from the well-studied lowland ecotype and the phenotypically
intermediate coastal ecotype (38).

Here, we test for GxWeather for two phenological timings by assigning
patterns of genetic effects on phenology at eight common gardens to many
patterns of weather covariance at these gardens. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development. We do this at eight common garden locations
spanning 17 degrees of latitude: these gardens cover the majority of the
latitudinal and climatic range of switchgrass and capture the most
comprehensive picture to date of the environmental variation this
species encounters. We define multiple ways phenological timings might
covary with weather (Table 1) and additional ways phenological timings
might vary by garden (SI Appendix, Section S1), then jointly re-estimate
genetic effects on these timings at all eight common gardens using the
set of these covariance matrices that significantly improved the modeled
log-likelihood when included (SI Appendix, Section S2-S5) (41). We use
the Bayesian framework \emph{mash} (multivariate adaptive shrinkage)
developed by (41), to refine effect size estimates from genome-wide
association (GWAS). \emph{Mash} allows us to identify and specify
multiple covariance structures among genetic effect estimates across
sites, including structures that represent covariance in weather
variables of interest. Importantly, this method circumvents statistical
biases in detecting genetic effects with the same or opposite signs
(42). To confirm our genetic mapping of GxWeather, we compare genomic
locations of the significant posterior effect estimates from \emph{mash}
to mapping results from an outbred mapping population grown at the same
sites. Our analyses allow us to describe the weather cues and types of
GxE affecting phenology in two divergent natural populations of
switchgrass across the species' latitudinal range.

@@

Genotypes from the Gulf and Midwest subpopulations had distinct
phenological timings and distinct patterns of phenological correlations
across our eight common garden sites (Figure~\ref{fig-map}). At the
three Texas common gardens (hereafter `Texas' gardens), located within
the natural range of the Gulf subpopulation, the onset of vegetative
growth, or `green-up date', for Gulf genotypes occurred before
Midwestern green-up, and the onset of reproductive growth, or `flowering
date', for Gulf genotypes occurred after Midwestern flowering
(Figure~\ref{fig-map} A). At the four northernmost common gardens
(hereafter `North' gardens), located within the natural range of the
Midwest subpopulation, both green-up and flowering of Gulf genotypes
occurred after that of Midwest genotypes. At the Oklahoma common garden,
located near the natural range limits of both the Gulf and the Midwest
subpopulations, Gulf and Midwest green-up occurred over the same period,
and Gulf genotype flowering occurred after Midwestern flowering
(Figure~\ref{fig-map} A). These patterns led to strong negative
phenotypic correlations for the onset of vegetative growth between the
North and Texas gardens, particularly in the Gulf and across all
individuals from the Gulf and Midwest subpopulations (hereafter, `Both'
subpopulations), and contributed to positive phenotypic correlations for
the onset of flowering that had larger magnitudes at more northern
gardens (Figure~\ref{fig-map} B).

Narrow-sense heritabilities (h\textsuperscript{2}) suggested that
rank-changing GxE for these phenotypes was present across the common
gardens (SI Appendix, Figure S1). h\textsuperscript{2} were typically
high at individual gardens: 59\% on average for green-up date, and 87\%
for flowering date. However, h\textsuperscript{2} were variable across
gardens, and green-up dates were uncorrelated (r\textsuperscript{2}
\textless{} 0.2) or negatively correlated between pairs of gardens
(Figure~\ref{fig-map} B). These negative and small correlations
undoubtedly contributed to the low h\textsuperscript{2} values for
green-up and flowering date when estimated jointly at all eight gardens:
h\textsuperscript{2} was 0.8\% for green-up and 23.2\% for flowering
date (SI Appendix, Figure S1).

@@

We expected patterns of GxE and GxWeather to vary at the locus,
subpopulation, and trait level; thus we were interested in modeling the
different ways effects might covary across our eight common gardens. We
jointly re-estimated genetic effects of a subset of SNPs across all
eight common gardens using \emph{mash} models, which can flexibly
capture many modes of covariance across gardens (SI Appendix, Sections
S1-S5, Datasets 1-6). To obtain genetic effects, we used subsets of
effects from garden- and subpopulation GWAS (SI Appendix, Section S2-S3;
Table S1). To obtain covariance structures, we specified both covariance
structures introduced in the initial \emph{mash} manuscript, and
GxWeather covariance matrices estimated from the covariance of empirical
weather patterns at each garden at specific times before the
phenological event (Table 1; SI Appendix, Section S1). Then, we used a
greedy algorithm to select covariance matrices from each model's set
that significantly improved the model likelihood (SI Appendix, Section
S4; Table S2). Finally, we fit \emph{mash} models using the selected set
of covariance matrices on a subset relatively unlinked
(r\textsuperscript{2} \textless{} 0.2), `strong' genetic effects with
low \emph{p}-values (SI Appendix, Section S5).

The loadings of genetic effects onto the GxE and GxWeather covariance
matrices included in each model provide information on genome-wide
patterns of SNP-environment interaction. For example, the phenotypic
correlations for the onset of vegetative growth had moderate negative
correlations between the Texas and North gardens, particularly in the
Gulf subpopulation and Both subpopulations (Figure~\ref{fig-map} B). If
these phenotypic correlations have a genetic component, they could be
partially or completely captured by the covariance structures specified
in the \emph{mash} model. In this case, the joint estimate of effects
for many SNPs could have high mixture proportions, or mass, on a
covariance matrix that captures the negative correlation between Texas
and North gardens. In our data, daylength 14 days prior to green-up date
has negative correlations between Texas and North gardens
(Figure~\ref{fig-covar} A); if many SNP effects have mass on this
GxWeather matrix, we would infer that the effects of those SNPs have
that form of GxWeather covariation.

@@

For green-up date, SNP-associated phenotypic effects covaried with
different weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the \emph{mash} model of Gulf green-up fell on a
covariance matrix constructed using the covariance of day length 14 days
prior to the date of green-up. The covariance matrix for this weather
cue was visually similar to the observed pattern of phenotypic
correlation for the timing of vegetative growth in the Gulf
subpopulation (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A).
\emph{Mash} models of the timing of vegetative growth in the Midwest
subpopulation and Both subpopulations did not include this GxWeather
covariance. The Midwest had 0.15\% weight on a GxWeather covariance
matrix of cumulative GDD for the 28d prior to green-up, a GxWeather
matrix that was visually similar to the Midwest's observed pattern of
phenotypic correlation (Figure~\ref{fig-map} B). Both subpopulations had
non-zero weights on two additional GxWeather covariance types, average
temperature one day prior to green-up, and the day length change in
seconds in the day prior to green-up (Figure~\ref{fig-covar} B). The
average temperature covariance matrix had negative covariances between
Texas and North gardens, though not as strong as the negative phenotypic
correlations seen in Both subpopulations (Figure~\ref{fig-map} B;
Figure~\ref{fig-covar} A). The Gulf subpopulation and Both
subpopulations had substantially more mass on GxWeather covariance
matrices than the Midwest population (Figure~\ref{fig-covar} C) for the
onset of vegetative growth.

@@

For green-up date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited rank-changing GxE, or a difference in sign
between pairs of Texas and North gardens (Figure~\ref{fig-effects} B; SI
Appendix, Fig S2). 78.7\% of pairwise comparisons between North and
Texas gardens had a difference in sign, while only 28.6\% and 0.2\% of
North-North or Texas-Texas comparisons had a difference in sign,
respectively (Figure~\ref{fig-effects} B). The majority of pairwise
effects for greenup for the Midwest (\textgreater73\%; 68-342 effects)
were indistinguishable, or had no GxE. The majority of pairwise effects
for Both (\textgreater85\%) subpopulations were the same sign, and
effects most often differed in magnitude between gardens within and
between the regions; however, these differences in effect sign and
magnitude were mostly between the MO site and other sites
(Figure~\ref{fig-effects} B; SI Appendix, Fig S2).

@@

We sought additional experimental support the occurrence and location of
our significant SNP effect re-estimates using an independent pseudo-F2
mapping population created from Gulf \& Midwest individuals and grown at
the same sites (Figure~\ref{fig-qtl} A,B). We conducted quantitative
trait loci (QTL) mapping of flowering as functions of five environmental
cues that we also used as covariance matrices in \emph{mash}, and
identified eight QTL for flowering date, eight QTL for flowering as a
function of day length change two days prior, one QTL for the start of
vegetative growth, and two QTL for vegetative growth as a function of
daylength change one day prior, all of which showed QTL by environment
interactions (SI Appendix, Figure S4). All QTL for flowering overlapped
one or more homologs from rice or \emph{A. thaliana} with functionally
validated roles in flowering (SI Appendix, Dataset 7); the QTL on Chr04K
overlapped a gene recently functionally validated for flowering in
switchgrass (43). All flowering and green-up QTL intervals contained at
least one SNP significant in at least one \emph{mash} run at a
log10-transformed Bayes Factor \textgreater{} 2, or in the 1\% tail of
significance, whichever was stricter (SI Appendix, Dataset 8). We also
looked for enrichments of \emph{mash} SNPs in the 1\% tail of
significance (the `\emph{mash} 1\% tail') within each QTL interval. At
the 5\% level, three QTL had enrichments of SNPs in the \emph{mash} 1\%
tail. Overall, there were five significant enrichments (p \textless{}
0.05, hypergeometric test) of SNPs in the \emph{mash} 1\% tail in the
QTL intervals. Thus, we were able to experimentally support the genomic
windows of some re-estimates of significant SNP effects with a QTL
mapping experiment using a separate mapping population.

@@

As the climate and the natural environment change, it is increasingly
critical to understand how patterns of plant-environment interactions
will change in response. To do this, we must understand the current
patterns of trait covariation across environments, the genetic
underpinnings of these patterns, and the cases where this covariation
can be altered. Here, we demonstrate that we can associate multiple
patterns of GxWeather with specific genomic regions using a switchgrass
diversity panel grown at eight common gardens. We can assign genetic
effects to both GxWeather patterns with interpretable weather-based
cues, and to unmeasured, site-based patterns. We use this approach to
study GxWeather for the timings of vegetative and reproductive
development in the deeply genetically diverged Gulf and Midwest
subpopulations of switchgrass.

Our analysis of the timing of vegetative and reproductive growth
revealed many alleles with rank-changing GxE, or sign changes in their
effects between the Texas and North gardens (Figure~\ref{fig-effects}
B). The Gulf \& Both subpopulations had rank-changing GxE for green-up
date, while the Midwest subpopulation had rank-changing GxE for
flowering date. As phenological timings are major components of plant
fitness, this result supports theoretical models that local adaptation
should involve antagonistic pleiotropy at the level of individual loci
(27--30). In addition, it advances local adaptation research by using
more than two field sites and a wide range of genetic variation to
determine that rank-changing GxE for phenological timings is common in
small genomic regions (15, 44, 45).

@@

\includegraphics{images/Figure_1_AB.png}

@@

\caption{\label{fig-effects}Types of GxE present between of pairs of
jointly re-estimated effects in eight common gardens, for effects with
lfsr \textless{} 0.05. a) Examples of effect patterns at three pairs of
sites with three types of GxE. All effects are for an alternate allele,
with the reference allele effect defined to be zero at both gardens and
represented by the dashed vertical line. Sign: Effects that differ in
sign at these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude:
Effects identical in sign (p \textless{} 0.05, lfsr) that differ in
magnitude by a factor of \textgreater0.4. Not Distinguishable: Effects
not distinguishable by magnitude nor sign of the effect, with no
measurable GxE. b) The fraction of effects with each GxE type for the
onset of vegetative growth (green-up date) and reproductive growth
(flowering date), within and between two genetic subpopulations. Common
gardens are grouped by the larger region they came from: North gardens
are within the natural range of the Midwest subpopulation, and include
MO, NE, MI, and SD, while Texas gardens are within the natural range of
the Gulf subpopulation, and include TX1, TX2, and TX3.}

@@

\acknow{We thank the Brackenridge Field laboratory, the Ladybird Johnson
Wildflower Center, and the Juenger laboratory for support with plant
care and propagation. This material is based upon work supported in part
by the Great Lakes Bioenergy Research Center, U.S. Department of Energy,
Office of Science, Office of Biological and Environmental Research under
Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of
Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, National
Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and
the Long-term Ecological Research Program (DEB 1832042) at the Kellogg
Biological Station.}

>>>>

\authorcontributions{T.E.J. designed research. D.B.L. contributed plant
material and resources. J.B., D.B.L., and T.E.J. designed and executed
field experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W.,
and T.E.J. hosted field experiments. A.H.M., L.Z., and S.A.S. conducted
statistical and computational analyses with input from T.E.J. and A.H.
The manuscript was written by A.H.M. with contributions from all
authors.}

@@

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depends on genetic effects (G), environmental
(e.g., weather) cues, and their interaction. Measuring GxWeather
interactions can aid prediction and manipulation of phenological
timings. Here, we map GxWeather effects on phenological timings in two
highly divergent switchgrass (\emph{Panicum virgatum}) populations using
repeated plantings of cloned individuals from these populations at eight
sites across the central United States. Some genetic effects covary with
interpretable weather-based cues. For example, in the Gulf population,
65\% of genetic effects on the timing of vegetative growth covary with
daylength 14 days prior to green-up date, and 33\% of genetic effects on
the timing of flowering covary with cumulative rainfall in the seven
days prior to flowering. However, most variation in genetic effects is
site-specific and cannot be attributed to variation in weather
variables. We demonstrate that we can identify genetic variation with
GxWeather and assign these loci to specific weather-based cues or other
patterns. We predict that selective breeding for particular alleles at
the GxWeather loci could alter flowering responsiveness in a photoperiod
or rainfall-specific way. More broadly, our approach proposes a refined
characterization of genotype-by-environment interactions in any species
phenotyped in multiple environments.
\end{abstract}

@@

Plant phenological traits are important components of plant fitness that
are affected by multiple external environmental cues (e.g.~degree of
winter chilling, day length, temperature, soil fertility, and water
availability), which signal existing or upcoming growing conditions
(1--4). Genetic responses to environmental cues determine the speed,
timing, and energy apportioned to vegetative and reproductive growth and
shape plant physiological responses, lifespan, and lifetime production
of viable seed. Day length (or photoperiod) is one of the most
predictable environmental cues, and genetic sensitivity to photoperiod
protects plants from potentially fatal consequences of phenological
responses to temperature cues at the ``wrong'' time of year. However,
the utility of specific environmental cues depends on both features of
the environment, such as cue predictability and relevance, and the
species' adaptive strategies (5). Species with wide natural
distributions can have multiple distinct environmentally cued
phenological responses. For example, populations of sunflower
(\emph{Helianthus annuus}) exhibit day-neutral, facultative short day,
and facultative long-day flowering responses, which vary with their
environments (6, 7). These distinct genetic responses in different
environments are known as genotype-by-environment interactions, or GxE.

@@

Previous research suggests that switchgrass phenological timings should
have GxWeather and that these timings could differ by genetic
subpopulation. Switchgrass is considered a short-day plant with
reproductive development strongly linked to day of the year (35).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (36, 37). We previously found that divergent
Midwest and Gulf genetic subpopulations of switchgrass have distinct
sets of environmental adaptations associated with each of two fitness
proxies, biomass and overwinter survival (38). The Midwest genetic
subpopulation is primarily composed of individuals from the well-studied
upland switchgrass ecotype (39, 40), while the Gulf subpopulation has
individuals from the well-studied lowland ecotype and the phenotypically
intermediate coastal ecotype (38).

Here, we test for GxWeather for the timing of two phenological traits by
loading patterns of genetic effects on phenology at eight common gardens
onto many patterns of weather covariance at these gardens. To do this,
we phenotyped a diversity panel of hundreds of switchgrass genotypes
from the Midwest and Gulf subpopulations for the timing of vegetative
and reproductive development. We did this at eight common garden
locations spanning 17 degrees of latitude: these gardens covered the
majority of the latitudinal and climatic range of switchgrass and
captured the most comprehensive picture to date of the environmental
variation this species encounters. We defined multiple ways phenological
traits might covary with weather (Table 1) and additional ways
phenological traits might vary by garden (SI Appendix, Section S1), then
jointly re-estimated genetic effects on these timings at all eight
common gardens using the set of these covariance matrices that
significantly improved the modeled log-likelihood when included (SI
Appendix, Section S2-S5) (41). We used the Bayesian framework
\emph{mash} (multivariate adaptive shrinkage) developed by (41), to
refine effect size estimates from genome-wide association (GWAS).
\emph{Mash} allowed us to identify and specify multiple covariance
structures among genetic effect estimates across sites, including
structures that represent covariance in weather variables of interest.
Importantly, this method circumvented statistical biases in detecting
genetic effects with the same or opposite signs (42). To confirm our
genetic mapping of GxWeather, we compared genomic locations of the
significant posterior effect estimates from \emph{mash} to mapping
results from an outbred mapping population grown at the same sites. Our
analyses allowed us to describe the weather cues and types of GxE
affecting phenology in two divergent natural populations of switchgrass
across the species' latitudinal range.

@@

Genotypes from the Gulf and Midwest subpopulations had distinct
phenological trait timings and distinct patterns of phenological trait
correlations across our eight common garden sites
(Figure~\ref{fig-map}). At the three Texas common gardens (hereafter
`Texas' gardens), located within the natural range of the Gulf
subpopulation, the onset of vegetative growth, or `green-up date', for
Gulf genotypes occurred before Midwestern green-up, and the onset of
reproductive growth, or `flowering date', for Gulf genotypes occurred
after Midwestern flowering (Figure~\ref{fig-map} A). At the four
northernmost common gardens (hereafter `North' gardens), located within
the natural range of the Midwest subpopulation, both green-up and
flowering of Gulf genotypes occurred after that of Midwest genotypes. At
the Oklahoma common garden, located near the natural range limits of
both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up
occurred over the same period, and Gulf genotype flowering occurred
after Midwestern flowering (Figure~\ref{fig-map} A). These patterns led
to strong negative phenotypic correlations for the onset of vegetative
growth between the North and Texas gardens, particularly in the Gulf and
in the population containing all individuals from the Gulf and Midwest
subpopulations (hereafter, `Both' subpopulations) and contributed to
positive phenotypic correlations for the onset of flowering that had
larger magnitudes at more northern gardens (Figure~\ref{fig-map} B).

Narrow-sense heritabilities (h\textsuperscript{2}) suggested that
rank-changing GxE for these phenotypes was present across the common
gardens (SI Appendix, Figure S1). h\textsuperscript{2} were typically
high at individual gardens: 59\% on average for green-up date, and 87\%
for flowering date. However, h\textsuperscript{2} was variable across
gardens, and green-up dates were uncorrelated (r\textsuperscript{2}
\textless{} 0.2) or negatively correlated between pairs of gardens
(Figure~\ref{fig-map} B). These negative and small correlations
undoubtedly contributed to the low h\textsuperscript{2} values for
green-up and flowering date when estimated jointly at all eight gardens:
h\textsuperscript{2} was 0.8\% for green-up and 23.2\% for flowering
date (SI Appendix, Figure S1).

@@

We expected patterns of GxE and GxWeather to vary at the locus,
subpopulation, and trait level. Thus, we were interested in modeling the
different ways effects might covary across our eight common gardens. We
jointly re-estimated the genetic effects of a subset of SNPs across all
eight common gardens using \emph{mash} models, which can flexibly
capture many modes of covariance across gardens (SI Appendix, Sections
S1-S5, Datasets 1-6). To obtain genetic effects, we used subsets of
effects from garden- and subpopulation GWAS (SI Appendix, Section S2-S3;
Table S1). To obtain covariance structures, we specified both covariance
structures introduced in the initial \emph{mash} manuscript, and
GxWeather covariance matrices estimated from the covariance of empirical
weather patterns at each garden at specific times before the
phenological event (Table 1; SI Appendix, Section S1). Then, we used a
greedy algorithm to select covariance matrices from each model's set
that significantly improved the model likelihood (SI Appendix, Section
S4; Table S2). Finally, we fit \emph{mash} models using the selected set
of covariance matrices on a subset relatively unlinked
(r\textsuperscript{2} \textless{} 0.2), `strong' genetic effects with
low \emph{p}-values (SI Appendix, Section S5).

The loadings of genetic effects onto the GxE and GxWeather covariance
matrices included in each model provide information on genome-wide
patterns of SNP-environment interaction. For example, the phenotypic
correlations for the onset of vegetative growth had moderate negative
correlations between the Texas and North gardens, particularly in the
Gulf subpopulation and Both subpopulations (Figure~\ref{fig-map} B). If
these phenotypic correlations have a genetic component, they could be
partially or completely captured by the covariance structures specified
in the \emph{mash} model. In this case, the joint estimate of effects
for many SNPs could have high mixture proportions, or mass, on a
covariance matrix that captures the negative correlation between Texas
(TX1, TX2, and TX3) and North (MO, NE, MI, and SD) gardens. In our data,
daylength 14 days prior to green-up date was negatively correlated
between Texas and North gardens (Figure~\ref{fig-covar} A); if many SNP
effects have mass on this GxWeather matrix, we would infer that the
effects of those SNPs have that form of GxWeather covariation.

@@

For green-up date, SNP-associated phenotypic effects covaried with
different weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the \emph{mash} model of Gulf green-up fell on a
covariance matrix constructed using the covariance of day length 14 days
prior to the date of green-up. The covariance matrix for this weather
cue was visually similar to the observed pattern of phenotypic
correlation for the timing of vegetative growth in the Gulf
subpopulation (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A).
\emph{Mash} models of the timing of vegetative growth in the Midwest
subpopulation and Both subpopulations did not include this GxWeather
covariance. The Midwest had 0.15\% weight on a GxWeather covariance
matrix of cumulative GDD for the 28d prior to green-up, a GxWeather
matrix that was visually similar to the Midwest's observed pattern of
phenotypic correlation (Figure~\ref{fig-map} B). Both subpopulations had
non-zero weights on two additional GxWeather covariance types, average
temperature one day prior to green-up, and the day length change in
seconds in the day prior to green-up (Figure~\ref{fig-covar} B). The
average temperature covariance matrix had negative covariances between
Texas and North gardens, though not as strong as the negative phenotypic
correlations observed in Both subpopulations (Figure~\ref{fig-map} B;
Figure~\ref{fig-covar} A). The Gulf subpopulation and Both
subpopulations had substantially more mass on GxWeather covariance
matrices than the Midwest population (Figure~\ref{fig-covar} C) for the
onset of vegetative growth.

@@

For green-up date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited rank-changing GxE, or a difference in sign
between pairs of Texas and North gardens (Figure~\ref{fig-effects} B; SI
Appendix, Fig S2). 78.7\% of pairwise comparisons between North and
Texas gardens had a difference in sign, while only 28.6\% and 0.2\% of
North-North or Texas-Texas comparisons had a difference in sign,
respectively (Figure~\ref{fig-effects} B). The majority of pairwise
effects for greenup for the Midwest (\textgreater73\%; 68-342 effects)
were indistinguishable, or had no GxE. The majority of pairwise effects
for Both subpopulations (\textgreater85\%) were the same sign, and
effects most often differed in magnitude between gardens within and
between the regions; however, these differences in effect sign and
magnitude were mostly between the MO site and other sites
(Figure~\ref{fig-effects} B; SI Appendix, Fig S2).

@@

We sought additional experimental support the occurrence and location of
our significant SNP effect re-estimates using an independent pseudo-F2
mapping population created from Gulf \& Midwest individuals and grown at
the same sites (Figure~\ref{fig-qtl} A,B). We conducted quantitative
trait loci (QTL) mapping of flowering as functions of five environmental
cues that we also used as covariance matrices in \emph{mash}, and
identified eight QTL for flowering date, eight QTL for flowering as a
function of day length change two days prior, one QTL for the start of
vegetative growth, and two QTL for vegetative growth as a function of
daylength change one day prior, all of which showed QTL by environment
interactions (SI Appendix, Figure S4). All QTL for flowering overlapped
with one or more homologs from rice or \emph{A. thaliana} with
functionally validated roles in flowering (SI Appendix, Dataset 7); the
QTL on Chr04K overlapped a gene recently functionally validated for
flowering in switchgrass (43). All flowering and green-up QTL intervals
contained at least one SNP significant in at least one \emph{mash} run
at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\% tail
of significance, whichever was stricter (SI Appendix, Dataset 8). We
also looked for enrichments of \emph{mash} SNPs in the 1\% tail of
significance (the `\emph{mash} 1\% tail') within each QTL interval. At
the 5\% level, three QTL had enrichments of SNPs in the \emph{mash} 1\%
tail. Overall, there were five significant enrichments (p \textless{}
0.05, hypergeometric test) of SNPs in the \emph{mash} 1\% tail in the
QTL intervals. Thus, we were able to experimentally support the genomic
windows of some re-estimates of significant SNP effects with a QTL
mapping experiment using a separate mapping population.

@@

As the climate and the natural environment change, it is increasingly
critical to understand how patterns of plant-environment interactions
will change in response. To do this, we must understand the current
patterns of trait covariation across environments, the genetic
underpinnings of these patterns, and the cases where this covariation
can be altered. Here, we demonstrate that we can associate multiple
patterns of GxWeather with specific genomic regions using a switchgrass
diversity panel grown at eight common gardens. We assigned genetic
effects to both GxWeather patterns with interpretable weather-based
cues, and to unmeasured, site-based patterns. We used this approach to
study GxWeather for the timings of vegetative and reproductive
development in the deeply genetically diverged Gulf and Midwest
subpopulations of switchgrass.

Our analysis of the timing of vegetative and reproductive growth
revealed many alleles with rank-changing GxE, or sign changes in their
effects between the Texas and North gardens (Figure~\ref{fig-effects}
B). The Gulf \& Both subpopulations had rank-changing GxE for green-up
date, while the Midwest subpopulation had rank-changing GxE for
flowering date. As phenological timings are major components of plant
fitness, this result supports theoretical models that local adaptation
should involve trade-offs at the level of individual loci (27--30). In
addition, it advances local adaptation research by using more than two
field sites and a wide range of genetic variation to determine that
rank-changing GxE for phenological traits is common in small genomic
regions (15, 44, 45).

@@

\includegraphics{images/Fig1_AB.png}

@@

\caption{\label{fig-effects}Types of GxE present between of pairs of
jointly re-estimated effects in eight common gardens, for effects with
lfsr \textless{} 0.05 at each pair of sites contrasted. a) Examples of
effect patterns at three pairs of sites with three types of GxE. All
effects are for an alternate allele, with the reference allele effect
defined to be zero at both gardens and represented by the dashed
vertical line. Sign: Effects that differ in sign at these pairs of
gardens (p \textless{} 0.05, lfsr). Magnitude: Effects identical in sign
(p \textless{} 0.05, lfsr) that differ in magnitude by a factor of
\textgreater0.4. Not Distinguishable: Effects not distinguishable by
magnitude nor sign of the effect, with no measurable GxE. b) The
fraction of effects with each GxE type for the onset of vegetative
growth (green-up date) and reproductive growth (flowering date), within
and between two genetic subpopulations. Common gardens are grouped by
the larger region they came from: North gardens are within the natural
range of the Midwest subpopulation, and include MO, NE, MI, and SD,
while Texas gardens are within the natural range of the Gulf
subpopulation, and include TX1, TX2, and TX3.}

@@

\acknow{We thank the Brackenridge Field laboratory, the Ladybird Johnson
Wildflower Center, and the Juenger laboratory for support with plant
care and propagation. This material is based upon work supported in part
by the Great Lakes Bioenergy Research Center, U.S. Department of Energy,
Office of Science, Office of Biological and Environmental Research under
Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of
Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, NIH grant
R35GM151108 and a Pew Scholarship to A.H., National Science Foundation
PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and the Long-term
Ecological Research Program (DEB 1832042) at the Kellogg Biological
Station.}
